---
title: PHQ-9
subtitle: Version 1.0.0
author:
- <h5 style="font-style:italic"> Tobias R. Spiller
- <h5 style="font-style:italic"> Or Duek
date: <h5 style="font-style:roman"> `r format(Sys.time(), '%d %B, %Y')`
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(readxl)

# Demographics
library(table1)
library(gtsummary)

# Analysis
#library("devtools")
#devtools::install_github("orduek/PsychPower")
library(PsychPower)
```

## 1. Import and prepare data

```{r PHQ-9 import data, message=FALSE, warning=TRUE, echo = FALSE}
# Import data
data_MDD_raw <- read_excel("/home/ptslab/Documents/powerLaw_newData/Ilan_PHQ9_and_GAD7_and_MH_Project_051223.xlsx",na = "NULL")
```

First, we select only one entry per individual.

```{r PHQ-9 filter, echo=TRUE, message=FALSE, warning=TRUE}
data_MDD_filtered <- data_MDD_raw %>%
  group_by(ID) %>%
  slice_head(n = 1) %>%
  ungroup()
```

Second, we define how the PHQ-9 items map onto the DSM-5 criteria.

```{r PHQ-9 map, echo=TRUE, message=FALSE, warning=TRUE}
data_MDD <- data_MDD_filtered %>%
  mutate(
    MDD_Symptom1 = as.numeric(PHQ9_2),
    MDD_Symptom2 = as.numeric(PHQ9_1),
    MDD_Symptom3 = as.numeric(PHQ9_5),
    MDD_Symptom4 = as.numeric(PHQ9_3),
    MDD_Symptom5 = as.numeric(PHQ9_8),
    MDD_Symptom6 = as.numeric(PHQ9_4),
    MDD_Symptom7 = as.numeric(PHQ9_6),
    MDD_Symptom8 = as.numeric(PHQ9_7),
    MDD_Symptom9 = as.numeric(PHQ9_9)) %>% 
  mutate(PHQ_9_Total = MDD_Symptom1 + MDD_Symptom2 + MDD_Symptom3 + MDD_Symptom4 +
          MDD_Symptom5 + MDD_Symptom6 + MDD_Symptom7 + MDD_Symptom8 + MDD_Symptom9) %>% 
  drop_na()
```

Third, we define the diagnosis.

```{r PHQ-9 Dx, echo=TRUE, message=FALSE, warning=TRUE}
data_MDD <- data_MDD %>%
  mutate(
    MDD_Diagnosis = ifelse(
      
      #Condition1
      (MDD_Symptom1 >= 2 | MDD_Symptom2 >= 2) & 
      
      #Condition2  
      rowSums(select(., starts_with("MDD_Symptom")) >= 2) >= 5,
      
      #Output
      TRUE, FALSE
    )
  )
```

Fourth, we compare demographics of individuals that meet and do not meet criteria.

```{r PHQ-9 compare demographics, message=FALSE, warning=TRUE, echo = FALSE}    
# Set Theme
theme_gtsummary_compact()

data_MDD$AGE <- as.numeric(data_MDD$AGE)

# Table 1
data_MDD %>%
  select(
    # Demographics
    AGE, GENDER, DEPRESSION_DX, MDD_Diagnosis, PHQ_9_Total) %>%
  tbl_summary(
    by = MDD_Diagnosis,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"),
    digits = ~ 2) %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "Met MDD criteria, No. (%) of Patients") %>%
  bold_labels()
```
<br><br>

Fifth, we subset the population that meets the DSM-5 criteria. [not shown]

```{r PHQ-9 subset, message=FALSE, warning=TRUE, echo = FALSE}  
# Subset
data_MDD_subset <- data_MDD %>% 
  filter(MDD_Diagnosis == TRUE)
```

```{r PHQ-9 further analysis, message=FALSE, warning=TRUE, echo = FALSE}
# Prepare dataset for further analysis // DATA1
data1 <- data_MDD_subset %>%  select (MDD_Symptom1:MDD_Symptom9)

## Extract variable names
variable_names <- c(names(data1), "total")

## Rename variables to "Q1:QN" for binarization
colnames(data1) <- c(paste0("Q", 1:(ncol(data1))))

## Add total score
data1 <- data1 %>% 
  mutate(total = rowSums(data1[1:ncol(data1)]))
```

## 2. Sample descriptive
#### 2.1 Demographics & charateristics
```{r PHQ-9 Table 1, echo=FALSE, message=FALSE}
# Table 1
data_MDD_subset %>%
  select(
    # Demographics
    AGE, GENDER, DEPRESSION_DX, MDD_Diagnosis, PHQ_9_Total) %>%
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})"),
    digits = ~ 2) %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_caption("Table 1. Patient Characteristics") %>%
  bold_labels()
``` 

#### 2.2 Cronbach's alpha
Selected sample
```{r PHQ-9 cronbach I, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(data1, select = (-total)))
cronbach$total
```
Full sample
```{r PHQ-9 cronbach II, echo=FALSE, message=FALSE}
data_all_cronbach <- data_MDD %>% select(MDD_Symptom1:MDD_Symptom9) 
cronbach_full <- psych::alpha(data_all_cronbach)
cronbach_full$total
```

Summary of items & histogram

```{r PHQ-9 summary, echo=FALSE, message=FALSE}
summary(data1)
hist(data1$total)
```

## 3. Binarizing & create datax 
Symptoms rated 2 or 3 will be considered as present. [not shown]
```{r PHQ-9 Binarizing, echo=FALSE, message=FALSE, warning=FALSE}
## set cut-off
cut_off <- 1 #will be used with <= // 

## Binarize
data1_binarized <- data1
for (i in 1:(ncol(data1)-1)){
  orig <- paste("q", i, sep = "")
  bin <- paste("Q", i, sep = "")
  data1_binarized[orig] <- dplyr::case_when(data1_binarized[bin]<= cut_off ~ 0, data1_binarized[bin]>cut_off ~ 1)  #0 = "Symptom absent", 1 = "Symptom present"

}

# Create new data frame
data2 <- data1_binarized %>% 
  select(total:ncol(data1_binarized)) %>% 
  select(-total) %>% 
  tibble()

## Count frequency of profiles
data2_counted <- plyr::count(data2[, ])

# Create sum score of endorsed symptoms
data2_counted <- data2_counted %>% 
  mutate(total_bin = rowSums(data2_counted)-freq)

## Create datax 
# Create full dataset
datax <- dplyr::left_join(data1_binarized, data2_counted)

## Create dataframes for PsychPower Package 
data_MDD_subset_selected <- data_MDD_subset %>%  select (MDD_Symptom1:MDD_Symptom9)
data_binarized <- PsychPower::binarize(data_MDD_subset_selected, cut_off = 1)

data_frequency <- PsychPower::pheno_frequency(
  data_binarized, target_columns = tidyselect::starts_with("v_bin"))
```

## 4. Description of combinations
Most common combination, its frequency & median frequency
```{r PHQ-9 Describe Number, echo=FALSE, message=FALSE}
desc_pheno <- PsychPower::describe_pheno(data_frequency, frequency = "freq") 
desc_pheno
```
10 Most common combinations
```{r PHQ-9 Describe 10 most, echo=FALSE, message=FALSE}
PsychPower::common_pheno(data_frequency, frequency = "freq", n_phenotypes = 10)
```

Combinations (N, %) reported by less or equal than in 1% of the individuals
```{r PHQ-9 Having less then 1 percent, echo=FALSE, message=FALSE}
one_percent_sample = round(nrow(data1)*0.01, 0)
data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()
round((data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()) / nrow(data2_counted), 3)
```

Individuals reporting one of the ten most common combinations
```{r PHQ-9 Having 10 most, echo=FALSE, message=FALSE}
data2_counted <- data2_counted %>% 
  arrange(desc(freq))
data2_counted[1:10,] %>% summarise(sum(freq))
```

Individuals reporting NOT one of the ten most common combinations
```{r PHQ-9  Not Having 10 most, echo=FALSE, message=FALSE}
data2_counted[11:nrow(data2_counted),] %>% summarise(sum(freq))
```

Individuals reporting 1% most common combinations
```{r PHQ-9 Describe 1 percent, echo=FALSE, message=FALSE}
one_percent =  round(nrow(data2_counted)*0.01, 0)

data2_counted[1:one_percent,] %>% summarise(sum(freq))
```

Individuals reporting one of the 50% least common combinations
```{r PHQ-9 Describe 50 percent, echo=FALSE, message=FALSE}
fifty_percent =  round(nrow(data2_counted)*0.5, 0)

data2_counted[fifty_percent:nrow(data2_counted),] %>% summarise(sum(freq))
```
Plot 100 most common combination
```{r Plot 100 most common, echo=FALSE, message=FALSE}
fig1 <- PsychPower::plot_pheno(data_frequency, frequency = "freq",
                               n_phenotypes = 100, color = "grey26")
fig1
```

## 5. Save & export data for Figures [not shown]
```{r Save & export, message=FALSE, include=FALSE}
#Export data for Figures
freq1_top_PHQ9 <- data2_counted %>% 
  top_n(freq, n = 100) %>% 
  select(freq)

### Figure 1
save(freq1_top_PHQ9, file = "../Results/freq1_top_PHQ9.RData")
```

## 6. Session Info
```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```
