---
title: PANSS - DSM-5
subtitle: Version 1.0.0
author:
- <h5 style="font-style:italic"> Tobias R. Spiller
- <h5 style="font-style:italic"> Or Duek
date: <h5 style="font-style:roman"> `r format(Sys.time(), '%d %B, %Y')`
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(readxl)

# Demographics
library(table1)
library(gtsummary)

# Analysis
#library("devtools")
#devtools::install_github("orduek/PsychPower")
library(PsychPower)
```

## 1. Import and prepare data

```{r PANSS import data, message=FALSE, warning=TRUE, echo = FALSE}
# Import data
df <- read.csv('/home/ptslab/Documents/powerLaw_newData/pnass01.csv')

df$Age <- as.numeric(df$interview_age) / 10 # change age variable accordingly

panssDf <- df[,213:242]
panssDf$subjectkey <- df$subjectkey
panssDf$age <- df$Age
panssDf$sex <- df$sex

### Clean dataset
panss <- panssDf[!duplicated(panssDf$subjectkey),]

# clean NAs
data_subset <- panss[,1:30]
panss <- panss[complete.cases(data_subset),] # only 4616 valid cases

data0 <- panss
data0$sex <- as.factor(data0$sex)
```

First, we define how the PANSS items map onto the DSM-5 criteria.

```{r PANSS map, echo=TRUE, message=FALSE, warning=TRUE}
# Algorithm
data_SZ_II <- data0 %>%
  mutate(
    # Criterion 1: delusions
    criterion_1 = pos_p1 >= 2,
    
    # Criterion 2: hallucinations
    criterion_2 = pos_p3 >= 2,
    
    # Criterion 3: disorganized speech
    criterion_3 = pos_p2 >= 2,
    
    # Criterion 4: grossly disorganized or catatonic behavior
    criterion_4 = (gps_g5 >= 2) | (gps_g7 >= 2),
    
    # Criterion 5: negative symptoms
    criterion_5 = (neg_n1 >= 2) | (neg_n2 >= 2) | (neg_n3 >= 2) | 
                  (neg_n4 >= 2) | (neg_n5 >= 2) | (neg_n6 >= 2) | 
                  (neg_n7 >= 2),
  ) %>% 
  mutate(PANSS_total_sum = rowSums(select(., pos_p1:gps_g16)),
         PANSS_positive_sum = rowSums(select(., pos_p1:pos_p7)),
         PANSS_negative_sum = rowSums(select(., neg_n1:neg_n7)),
         PANSS_global_sum = rowSums(select(., gps_g1:gps_g16))
  )

```

Second, we define the diagnosis

```{r PANSS Dx, echo=TRUE, message=FALSE, warning=TRUE}
data_SZ_II <- data_SZ_II %>%
  mutate(
    # Defin Dx
    PANSS_SZ_II_DX = (
      # Condition 1: At least one of the criterion 1-3 is TRUE
      (criterion_1 | criterion_2 | criterion_3) &
      
      # Condition 2: At least two of the criterion 1-5 are TRUE
      (sum(criterion_1, criterion_2, criterion_3, criterion_4, criterion_5) >= 2)
    )
  )
```

Third, we compare demographics of individuals that meet and do not meet criteria.

```{r PANSS compare demographics, message=FALSE, warning=TRUE, echo = FALSE}    
# Set Theme
theme_gtsummary_compact()

# Table 1
data_SZ_II %>%
  select(
    # Demographics
    age, sex, PANSS_SZ_II_DX,  
    PANSS_total_sum, PANSS_positive_sum, PANSS_negative_sum, PANSS_global_sum) %>%
  tbl_summary(
    by = PANSS_SZ_II_DX,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    digits = ~ 2) %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "Met SZ_II criteria, No. (%) of Patients") %>%
  bold_labels()
```
<br><br>

Fourth, we subset the population that meets the DSM-5 criteria. [not shown]

```{r PANSS subset, message=FALSE, warning=TRUE, echo = FALSE}  
# Subset
data_SZ_II_subset <- data_SZ_II %>% 
  filter(PANSS_SZ_II_DX == TRUE)
```

```{r PANSS further analysis, message=FALSE, warning=TRUE, echo = FALSE}
# Prepare dataset for further analysis // DATA1
data1 <- data_SZ_II_subset %>%  select (criterion_1:criterion_5)

data1 <- as.data.frame(lapply(data1, as.numeric))
  
## Extract variable names
variable_names <- c(names(data1), "total")

## Rename variables to "Q1:QN" for binarization
colnames(data1) <- c(paste0("Q", 1:(ncol(data1))))

## Add total score
data1 <- data1 %>% 
  mutate(total = rowSums(data1[1:ncol(data1)]))
```

## 2. Sample Descriptive
#### 2.1 Demographics & charateristics
```{r PANSS Table 1, echo=FALSE, message=FALSE}
# Table 1
data_SZ_II_subset %>%
  select(
    # Demographics
    age, sex, PANSS_SZ_II_DX,  
    PANSS_total_sum, PANSS_positive_sum, PANSS_negative_sum, PANSS_global_sum) %>%
  tbl_summary(
     statistic = list(
      all_continuous() ~ "{mean} ({sd})"),
    digits = ~ 2) %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_caption("Table 1. Patient Characteristics") %>%
  bold_labels()
``` 

#### 2.2 Cronbach's alpha
Selected sample
```{r PANSS cronbach I, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(data1, select = (-total)))
cronbach$total
```
Full sample
```{r PANSS cronbach II, echo=FALSE, message=FALSE}
data_cronbach <- data0 %>% select(pos_p1:gps_g16)
cronbach_full <- psych::alpha(data_cronbach)
cronbach_full$total
```

Summary of items & histogram
```{r PANSS summary, echo=FALSE, message=FALSE}
summary(data1)
hist(data1$total)
```

## 3. Binarizing & create datax 
Symptoms rated 3-7 will be considered as present.[not shown]
```{r PANSS Binarizing, echo=FALSE, message=FALSE, warning=FALSE}
## set cut-off
cut_off <- 0 #will be used with <= // 

## Binarize
data1_binarized <- data1
for (i in 1:(ncol(data1)-1)){
  orig <- paste("q", i, sep = "")
  bin <- paste("Q", i, sep = "")
  data1_binarized[orig] <- dplyr::case_when(data1_binarized[bin]<= cut_off ~ 0, data1_binarized[bin]>cut_off ~ 1)  #0 = "Symptom absent", 1 = "Symptom present"

}

# Create new data frame
data2 <- data1_binarized %>% 
  select(total:ncol(data1_binarized)) %>% 
  select(-total) %>% 
  tibble()


## Count frequency of profiles
data2_counted <- plyr::count(data2[, ]) 

# Create sum score of endorsed symptoms
data2_counted <- data2_counted %>% 
  mutate(total_bin = rowSums(data2_counted)-freq)


## Create datax 
# Create full dataset
datax <- dplyr::left_join(data1_binarized, data2_counted)


## Create dataframes for PsychPower Package 
data_SZ_II_subset_selected <- data1[,1:5]

data_binarized <- PsychPower::binarize(data_SZ_II_subset_selected, cut_off = 0)

data_frequency <- PsychPower::pheno_frequency(
  data_binarized, target_columns = tidyselect::starts_with("v_bin"))
```


## 4. Description of combinations
Most common combination, its frequency & median frequency
```{r PANSS Describe Number, echo=FALSE, message=FALSE}
desc_pheno <- PsychPower::describe_pheno(data_frequency, frequency = "freq") 
desc_pheno
```

10 Most common combinations
```{r PANSS Describe 10 most, echo=FALSE, message=FALSE}
PsychPower::common_pheno(data_frequency, frequency = "freq", n_phenotypes = 10)
```

Combinations (N, %) reported by less or equal than in 1% of the individuals
```{r PANSS Having less then 1 percent, echo=FALSE, message=FALSE}
one_percent_sample = round(nrow(data1)*0.01, 0)
data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()

round((data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()) / nrow(data2_counted), 3)
```

Individuals reporting one of the ten most common combinations
```{r PANSS Having 10 most, echo=FALSE, message=FALSE}
data2_counted <- data2_counted %>% 
  arrange(desc(freq))
data2_counted[1:10,] %>% summarise(sum(freq))
```

Individuals reporting NOT one of the ten most common combinations
```{r PANSS  Not Having 10 most, echo=FALSE, message=FALSE}
data2_counted[11:nrow(data2_counted),] %>% summarise(sum(freq))
```

Individuals reporting 1% most common combinations
```{r PANSS Describe 1 percent, echo=FALSE, message=FALSE}
one_percent =  round(nrow(data2_counted)*0.01, 0)

data2_counted[1:one_percent,] %>% summarise(sum(freq))
```

Individuals reporting one of the 50% least common combinations
```{r PANSS Describe 50 percent, echo=FALSE, message=FALSE}
fifty_percent =  round(nrow(data2_counted)*0.5, 0)

data2_counted[fifty_percent:nrow(data2_counted),] %>% summarise(sum(freq))
```

Plot 100 most common combination
```{r Plot 100 most common, echo=FALSE, message=FALSE}
fig1 <- PsychPower::plot_pheno(data_frequency, frequency = "freq",
                               n_phenotypes = 100, color = "grey26")
fig1
```

## 6. Save & export data for Figures [not shown]
```{r Save & export, message=FALSE, include=FALSE}
### Figure 1a
freq1_top_Sz_DSM <- data2_counted %>% 
  top_n(freq, n = 100) %>% 
  select(freq)
  
save(freq1_top_Sz_DSM, file = "../Results/freq1_top_Sz_DSM.RData")
```

## 7. Session Info
```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```
