---
title: PTSD
subtitle: Version 0.1.0
author:
- <h5 style="font-style:italic"> Tobias R. Spiller
- <h5 style="font-style:italic"> Or Duek
date: <h5 style="font-style:roman"> `r format(Sys.time(), '%d %B, %Y')`
output:
  html_document:
    toc: yes
    toc_depth: 5
    toc_float: yes
---

```{r Setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = TRUE,
	warning = TRUE
)
```

```{r Load Libraries, message=FALSE, warning=TRUE, include=FALSE}
# Data handling
library(tidyverse)
library(readxl)

# Demographics
library(table1)
library(gtsummary)

# Additional
library(foreign)
library(ReIns)
library(psych)
library(poweRlaw)
```

## 1. Import and prepare data

```{r PCL import data, message=FALSE, warning=TRUE, echo = FALSE}
# Import data
dfPCL5 <- read.spss('/home/ptslab/Documents/powerLaw_newData/PCLpct19.sav', to.data.frame = TRUE)

#### Prepare dataset
data_subset <- dfPCL5[,5:24]

# omitting data just based on PCL NAs
dfPCL5 <- dfPCL5[complete.cases(data_subset), ]
dfPCL5$totalPCL <- dfPCL5$PCLN01 + dfPCL5$PCLN02 + dfPCL5$PCLN03 + dfPCL5$PCLN04 + dfPCL5$PCLN05 + 
  dfPCL5$PCLN06 + dfPCL5$PCLN07 + dfPCL5$PCLN08 + dfPCL5$PCLN09 + dfPCL5$PCLN10 + dfPCL5$PCLN11 +
  dfPCL5$PCLN12 + dfPCL5$PCLN13 + dfPCL5$PCLN14 + dfPCL5$PCLN15 + dfPCL5$PCLN16 + dfPCL5$PCLN17 +
  dfPCL5$PCLN18 + dfPCL5$PCLN19 + dfPCL5$PCLN20

###### 2.1 Select variables for methods and materials // DATA0
data0 <- dfPCL5 %>% # xxxxx = dataset after cleaning before selection
  select("AGE_OCT01", "FEMALE", "totalPCL") #include age, sex, variables of interest
 
data0$FEMALE <- as.factor(data0$FEMALE)
```

First, we define the diagnosis

```{r PCL Dx, echo=TRUE, message=FALSE, warning=TRUE}
data_PTSD <- dfPCL5 %>%
  mutate(
    PTSD_Diagnosis = ifelse(
      (PCLN01 > 2 | PCLN02 > 2 | PCLN03 > 2 | PCLN04 > 2 | PCLN05 > 2) &
      (PCLN06 > 2 | PCLN07 > 2) &
      (PCLN08 > 2 | PCLN09 > 2 | PCLN10 > 2 | PCLN11 > 2 | PCLN12 > 2 | PCLN13 > 2 | PCLN14 > 2) &
      (PCLN15 > 2 | PCLN16 > 2 | PCLN17 > 2 | PCLN18 > 2 | PCLN19 > 2 | PCLN20 > 2),
      TRUE, FALSE
    )
  )
```

Second, we compare demographics of individuals that meet and do not meet criteria.

```{r PCL compare demographics, message=FALSE, warning=TRUE, echo = FALSE}    
# Add labels
# Set Theme
theme_gtsummary_journal(journal = "jama")
theme_gtsummary_compact()

# Table 1
data_PTSD %>%
  select(
    # Demographics
    AGE_OCT01, FEMALE, PTSD_Diagnosis, totalPCL) %>%
  tbl_summary(
    by = PTSD_Diagnosis,
    statistic = list(
      all_continuous() ~ "{mean} ({sd})")) %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_spanning_header(c("stat_1", "stat_2") ~ "Met PTSD criteria, No. (%) of Patients") %>%
  bold_labels() 
```
<br><br>
Third, we subset the population that meets the DSM-5 criteria. [not shown]

```{r PCL subset, message=FALSE, warning=TRUE, echo = FALSE}  
# Subset
data_PTSD_subset <- data_PTSD %>% 
  filter(PTSD_Diagnosis == TRUE)
```

```{r PCL further analysis, message=FALSE, warning=TRUE, echo = FALSE}
# Prepare dataset for further analysis // DATA1
data1 <- data_PTSD_subset %>%  select (PCLN01:PCLN20)

## Extract variable names
variable_names <- c(names(data1), "total")

## Rename variables to "Q1:QN" for binarization
colnames(data1) <- c(paste0("Q", 1:(ncol(data1))))

## Add total score
data1 <- data1 %>% 
  mutate(total = rowSums(data1[1:ncol(data1)]))
```

## 2. Sample Descriptive
```{r PCL Table 1, echo=FALSE, message=FALSE}
# Table 1
data_PTSD_subset %>%
  select(
    # Demographics
     AGE_OCT01, FEMALE, PTSD_Diagnosis, totalPCL) %>%
  tbl_summary(
    statistic = list(
      all_continuous() ~ "{mean} ({sd})")) %>% 
  modify_header(label ~ "**Variable**") %>%
  modify_caption("Table 1. Patient Characteristics") %>%
  bold_labels()
``` 

Cronbach's alpha
```{r PCL cronbach, echo=FALSE, message=FALSE}
cronbach <- psych::alpha(subset(data1, select = (-total)))
cronbach$total
```

Summary of items & histogram
```{r PCL summary, echo=FALSE, message=FALSE}
summary(data1)
hist(data1$total)
```
## 2.1 Correlation matrix
```{r}
dataCor <- data1[,1:20]
correlations <- rcorr(as.matrix(dataCor))
# Assuming data0 is your data frame
# Create a vector with the correct order of column names
correct_order <- paste0("Q", 1:20)

# Create a mask for significant correlations
cor_mat <- correlations$r  # Correlation matrix
p_mat <- correlations$P  # P-value matrix
sig_level <- 0.05  # Significance level
mask <- p_mat <= sig_level  # Matrix to mask non-significant correlations

# Plot using corrplot
corrplot::corrplot(cor_mat, type = "upper", order = "original", 
                    sig.level = sig_level, p.mat=p_mat,
                   insig = "blank", addCoef.col = "black", # Show correlation coefficient
                   tl.col="black", tl.srt=45) # Text label color and rotation


```



## 3. Binarizing & create datax 
Symptoms rated 2 or 3 will be considered as present.[not shown]
```{r PCL Binarizing, echo=FALSE, message=FALSE, warning=FALSE}
## set cut-off
cut_off <- 1 #will be used with <= // 

## Binarize
data1_binarized <- data1
for (i in 1:(ncol(data1)-1)){
  orig <- paste("q", i, sep = "")
  bin <- paste("Q", i, sep = "")
  data1_binarized[orig] <- dplyr::case_when(data1_binarized[bin]<= cut_off ~ 0, data1_binarized[bin]>cut_off ~ 1)  #0 = "Symptom absent", 1 = "Symptom present"

}

# Create new data frame
data2 <- data1_binarized %>% 
  select(total:ncol(data1_binarized)) %>% 
  select(-total) %>% 
  tibble()


## Count frequency of profiles
data2_counted <- plyr::count(data2[, ]) # 510

# Create sum score of endorsed symptoms
data2_counted <- data2_counted %>% 
  mutate(total_bin = rowSums(data2_counted)-freq)


## Create datax 
# Create full dataset
datax <- dplyr::left_join(data1_binarized, data2_counted)


## Create dataframes for PsychPower Package 
data_PTSD_subset_selected <- data_PTSD_subset %>%  select (PCLN01:PCLN20)
data_binarized <- PsychPower::binarize(data_PTSD_subset_selected, cut_off = 1)

data_frequency <- PsychPower::pheno_frequency(
  data_binarized, target_columns = tidyselect::starts_with("v_bin"))
```

```{r PCL Save data I, echo=FALSE, message=FALSE}
# Save for further analysis
write_csv2(data1, "PCL/Generated_data/binarized.csv")
write_csv2(data2_counted, "PCL/Generated_data/Binarized_freq_count.csv")
write_csv2(datax, "PCL/Generated_data/Binarized_Matched_freq_count.csv")
```

## 4. Description of Profiles
Severity
```{r PCL Describe Severity, echo=FALSE, message=FALSE}
# Summed severity
summary(datax$total)
hist(datax$total)
```

Most common Phenotype
```{r PCL Describe Number, echo=FALSE, message=FALSE}
desc_pheno <- PsychPower::describe_pheno(data_frequency, frequency = "freq") 
desc_pheno
```

10 Most common Phenotype
```{r PCL Describe 10 most, echo=FALSE, message=FALSE}
PsychPower::common_pheno(data_frequency, frequency = "freq", n_phenotypes = 10)
```


Profiles reported less or equal to 5 time
```{r PCL Describe 5 or less, echo=FALSE, message=FALSE}
data2_counted %>% filter(freq<=5) %>% nrow() # 
```

Profiles reported once
```{r PCL Describe once, echo=FALSE, message=FALSE}
data2_counted %>% filter(freq==1) %>% nrow() # 
```

Profiles (N, %) reported by less or equal than in 1% of the individuals

```{r PCL Having less then 1 percent, echo=FALSE, message=FALSE}
one_percent_sample = round(nrow(data1)*0.01, 0)
data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()

round((data2_counted %>% filter(freq<=one_percent_sample) %>% nrow()) / nrow(data2_counted), 3)
```

individuals reporting one of the ten most common profiles
```{r PCL Having 10 most, echo=FALSE, message=FALSE}
data2_counted <- data2_counted %>% 
  arrange(desc(freq))
data2_counted[1:10,] %>% summarise(sum(freq))
```

individuals reporting NOT one of the ten most common profiles
```{r PCL  Not Having 10 most, echo=FALSE, message=FALSE}
data2_counted[11:nrow(data2_counted),] %>% summarise(sum(freq))
```

individuals reporting 1% Most common Profiles
```{r PCL Describe 1 percent, echo=FALSE, message=FALSE}
one_percent =  round(nrow(data2_counted)*0.01, 0)

data2_counted[1:one_percent,] %>% summarise(sum(freq))
```

individuals reporting one of the 50% least common profiles
```{r PCL Describe 50 percent, echo=FALSE, message=FALSE}
fifty_percent =  round(nrow(data2_counted)*0.5, 0)

data2_counted[fifty_percent:nrow(data2_counted),] %>% summarise(sum(freq))
```


Plot 100 most common profiles
```{r Plot 100 most common, echo=FALSE, message=FALSE}
fig1 <- PsychPower::plot_pheno(data_frequency, frequency = "freq",
                               n_phenotypes = 100, color = "grey26")
fig1

# save
pdf("Images/Top_100_Phenotypes_PCL.pdf", width=8, height=8)
fig1
dev.off() 

# FOR Combined figure
freq1_top  <- data2_counted %>% 
  top_n(freq, n = 100) %>% 
  select(freq)
```


## 5. Assess & compare distributions 
Test
```{r Test distributions, echo=FALSE, message=FALSE}
#freq_distribution <- PsychPower::pheno_distributions(data_frequency, frequency = "freq")

#set.seed(241)

#desc_freq <- PsychPower::describe_pheno_distr(freq_distribution, bootStrap = T,
                          #        nBoots = 5, nCores = 2)
#round(desc_freq, 3)
```

Compare distributions

```{r Compare distributions, echo=FALSE, message=FALSE}
#comp_fit_freq <- PsychPower::compare_pheno_distr(freq_distribution)

#comp_fit_freq
```

## 6. Save & export data for Figures [not shown]
```{r Save & export, message=FALSE, include=FALSE}
#Export data for Figures 
### Figure 1a
freq1_top_PCL <- freq1_top 
save(freq1_top_PCL, file = "PCL/Generated_data/freq1_top_PCL.RData")

### Figure 1b
# Export m_pl & m_ln
#res_pl_PCL <- plot(freq_distribution[[1]])
#line_pl_PCL <- lines(freq_distribution[[1]])
#line_ln_PCL <- lines(freq_distribution[[2]])
#line_ex_PCL <- lines(freq_distribution[[3]])

# save(res_pl_PCL, file = "PCL/Generated_data/res_pl_PCL.RData")
# save(line_pl_PCL, file = "Analysis/PCL/Generated Data/line_pl_PCL.RData")
# save(line_ln_PCL, file = "Analysis/PCL/Generated Data/line_ln_PCL.RData")
# save(line_ex_PCL, file = "Analysis/PCL/Generated Data/line_ex_PCL.RData")
```

## 7. Session Info
```{r Session Info, echo=FALSE, message=FALSE}
sessionInfo()
```
